---
title: "Ha_EDA2"
author: "Thu Ha Nguyen"
date: "10/28/2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Variables used:
- unemployment rate
- # of households (maybe not)
- median household income
- percent poor household (<$10,000)
- bachelor degree

Variables used if presidential year included:
- employment rate (with interaction)
- unemployment rate 
- in labor force (with interaction)
- # of household (with interaction)
- median household income (with interaction)
- percent poor household (<$10,000) (with interaction)
- bachelor degree (with interaction)

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
```


```{r}
final_county <- read_csv("~/jvstats/final_county.csv")
nrow(final_county)
ncol(final_county)

simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}


final_county_tableau <- final_county %>%
  mutate(County = simpleCap(CTYNAME),
         State = "Minnesota")

#write_excel_csv(final_county_tableau, "~/jvstats/tableau_county.csv")

fips.df <- read_excel("~/jvstats/NEW_DATA/all-geocodes-v2016.xlsx", skip = 4, col_names = TRUE)
head(fips.df)
fips.df <- fips.df %>%
  filter(`State Code (FIPS)` == "27") %>%
  filter(`County Code (FIPS)` != "000") %>%
  filter(str_sub(`Area Name (including legal/statistical area description)`, -6, -1) == "County") %>%
  mutate(CTYNAME = str_to_lower(str_sub(`Area Name (including legal/statistical area description)`, 1, str_length(`Area Name (including legal/statistical area description)`) - 7)))
head(fips.df)
View(fips.df)
#write_excel_csv(fips.df, "~/jvstats/NEW_DATA/EXPORTED_DATA/fips.csv")
nrow(fips.df)
```

#EDA voterpercent
```{r}
final_county %>%
  ggplot(aes(x=VOTEPERCENT)) +
    geom_histogram() + facet_wrap(~MIDTERM) 

final_county %>%
  ggplot(aes(x=VOTEPERCENT)) +
    geom_histogram() + facet_wrap(~YEAR) 
```

#visualization for congressional district
```{r}
turnout_pct <- read_csv("~/jvstats/NEW_DATA/EXPORTED_DATA/turnout_pct.csv")
turnout_district <- turnout_pct %>%
  filter(!is.na(TOTVOTERS)) %>%
  group_by(CONGDIST, YEAR) %>%
  summarize(VOTE = sum(TOTVOTERS))
head(turnout_pct)
ggplot(data=turnout_district, aes(x=YEAR, y=VOTE, color=as.factor(CONGDIST))) +
    geom_line()
# Congressional districts have relatively equal population. The persistent difference in voter turnout suggests some persistent difference in socio-economic-political factors in these districts. 
ggplot(data=final_county, aes(x=YEAR, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_line()

```
#EDA with occupation
```{r}

ggplot(data=final_county, aes(x=`OCC.Agriculture`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +geom_point() + facet_wrap(~YEAR) +xlab("Agriculture")
ggplot(data=final_county, aes(x=`OCC.Agriculture`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Agriculture")

summary(lm(VOTEPERCENT~as.numeric(`OCC.Agriculture`), data = final_county))
summary(lm(VOTEPERCENT~as.numeric(`OCC.Agriculture`), data=subset(final_county,MIDTERM==1)))

```


```{r}
ggplot(data=final_county, aes(x=`OCC.Transportation`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +geom_point() + facet_wrap(~YEAR) +xlab("transportation")
ggplot(data=final_county, aes(x=`OCC.Transportation`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("transportation")

summary(lm(VOTEPERCENT~OCC.Transportation, data = final_county))
summary(lm(VOTEPERCENT~OCC.Transportation, data=subset(final_county,MIDTERM==1)))

```

```{r}
ggplot(data=final_county, aes(x=`OCC.WholesaleTrade`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +geom_point() + facet_wrap(~YEAR) +xlab("transportation")
ggplot(data=final_county, aes(x=`OCC.WholesaleTrade`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("transportation")

summary(lm(VOTEPERCENT~OCC.WholesaleTrade, data = final_county))
summary(lm(VOTEPERCENT~OCC.WholesaleTrade, data=subset(final_county,MIDTERM==1)))

```

#EDA with employment
```{r}
#final_county$EMPLOY.HC03_EST_VC01 guessing this is percentage?
ggplot(data=final_county, aes(x=`EMPLOY.HC03_EST_VC01`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Employed; Estimate; Population 16 years and over")
ggplot(data=final_county, aes(x=`EMPLOY.HC03_EST_VC01`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Employed; Estimate; Population 16 years and over")
#comment: unclear relationship for midterm election, negative rela for presidential election
#explanation: maybe unemployment policy doesn't matter in midterm election

summary(lm(VOTEPERCENT~`EMPLOY.HC03_EST_VC01`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`EMPLOY.HC03_EST_VC01`, data=subset(final_county,MIDTERM==1)))

```

```{r}
ggplot(data=final_county, aes(x=`EMPLOY.HC04_EST_VC01`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Unemployed; Estimate; Population 16 years and over")
ggplot(data=final_county, aes(x=`EMPLOY.HC04_EST_VC01`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Unemployed; Estimate; Population 16 years and over")
#comment:slight negative rela for midterm election, more negative rela for presidential one

summary(lm(VOTEPERCENT~`EMPLOY.HC04_EST_VC01`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`EMPLOY.HC04_EST_VC01`, data=subset(final_county,MIDTERM==1)))
```

```{r}
#final_county$EMPLOY.HC02_EST_VC01 this is also rate right?
ggplot(data=final_county, aes(x=`EMPLOY.HC02_EST_VC01`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("In labor force; Estimate; Population 16 years and over")
ggplot(data=final_county, aes(x=`EMPLOY.HC02_EST_VC01`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("In labor force; Estimate; Population 16 years and over")
#comment: seems to be no relationship for midterm election, but positive rela for presidential one

summary(lm(VOTEPERCENT~`EMPLOY.HC02_EST_VC01`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`EMPLOY.HC02_EST_VC01`, data=subset(final_county,MIDTERM==1)))
```



#EDA with income
- question: what to use: households, families?
```{r}
#final_county$INCOME.HC01_EST_VC01 don't understand what this variable mean, defintely not rate
#final_county$INCOME.HC01_EST_VC01
ggplot(data=final_county, aes(x=log(`INCOME.HC01_EST_VC01`), y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Households; Estimate; Total (log scale)")
ggplot(data=final_county, aes(x=log(`INCOME.HC01_EST_VC01`), y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Households; Estimate; Total (log scale)")

summary(lm(VOTEPERCENT~`INCOME.HC01_EST_VC01`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`INCOME.HC01_EST_VC01`, data=subset(final_county,MIDTERM==1)))
```


```{r}
#final_county$INCOME.HC01_EST_VC13
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC13, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Households; Estimate; Median income (dollars)")
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC13, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Households; Estimate; Median income (dollars)")
#comment:positive relationship

summary(lm(VOTEPERCENT~INCOME.HC01_EST_VC13*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`INCOME.HC01_EST_VC13`, data=subset(final_county,MIDTERM==1)))
```

```{r}
#final_county$INCOME.HC01_EST_VC02 this is rate right?
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC02, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Households; Estimate; Less than $10,000")
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC02, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Households; Estimate; Less than $10,000")
#comment: negative relation between votepercent and poor households (does < $10,000 count as poor?)

summary(lm(VOTEPERCENT~INCOME.HC01_EST_VC02*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`INCOME.HC01_EST_VC02`, data=subset(final_county,MIDTERM==1)))
```
##don't understand what this variable means
```{r}
#final_county$INCOME.HC01_EST_VC18 What does this variable mean?
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC18, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Households; Estimate; PERCENT IMPUTED - Household income in the past 12 months")
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC18, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Households; Estimate; PERCENT IMPUTED - Household income in the past 12 months")
#comment: negative relationship

summary(lm(VOTEPERCENT~INCOME.HC01_EST_VC18*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`INCOME.HC01_EST_VC18`, data=subset(final_county,MIDTERM==1)))
```

#EDA with education
```{r}
ggplot(data=final_county, aes(x=`BACHELOR'S DEGREE >25`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Bachelor Degree >25")
ggplot(data=final_county, aes(x=`BACHELOR'S DEGREE >25`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Bachelor degree >25")
#comment: positive rela

summary(lm(VOTEPERCENT~`BACHELOR'S DEGREE >25`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`BACHELOR'S DEGREE >25`, data=subset(final_county,MIDTERM==1)))
```

```{r}
ggplot(data=final_county, aes(x=`BACHELOR'S DEGREE`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Bachelor Degree")
ggplot(data=final_county, aes(x=`BACHELOR'S DEGREE`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Bachelor degree")
#comment: positive rela

summary(lm(VOTEPERCENT~`BACHELOR'S DEGREE`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`BACHELOR'S DEGREE`, data=subset(final_county,MIDTERM==1)))
```
```{r}
ggplot(data=final_county, aes(x=`HIGHSCHOOL DEGREE`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("High school Degree")
ggplot(data=final_county, aes(x=`HIGHSCHOOL DEGREE`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("High school degree")
#comment: negative rela for presidential year, not clear rela for midterm year

summary(lm(VOTEPERCENT~`HIGHSCHOOL DEGREE`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`HIGHSCHOOL DEGREE`, data=subset(final_county,MIDTERM==1)))
```


```{r}
ggplot(data=final_county, aes(x=`HIGHSCHOOL DEGREE >25`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("High school Degree > 25")
ggplot(data=final_county, aes(x=`HIGHSCHOOL DEGREE >25`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("High school degree > 25")
#comment: negative rela for presidential year, not clear rela for midterm year

summary(lm(VOTEPERCENT~`HIGHSCHOOL DEGREE >25`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`HIGHSCHOOL DEGREE >25`, data=subset(final_county,MIDTERM==1)))

summary(lm(VOTEPERCENT~`HIGHSCHOOL DEGREE` + `BACHELOR'S DEGREE` + `GRADUATE DEGREE >25`, data=subset(final_county,MIDTERM==1)))

```
```{r}
ggplot(data=final_county, aes(x=`GRADUATE DEGREE >25`, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Graduate Degree > 25")
ggplot(data=final_county, aes(x=`GRADUATE DEGREE >25`, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Graduate degree > 25")
#comment: positive rela for presidential year, but not clear rela for midterm year (this is surprising?)

summary(lm(VOTEPERCENT~`GRADUATE DEGREE >25`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`GRADUATE DEGREE >25`, data=subset(final_county,MIDTERM==1)))
```


#Sexage
```{r}
final_county <- final_county %>%
  mutate(PercentMale = TOTALMALE/ POPULATION, 
         PercentFemale = TOTALFEMALE/ POPULATION
         )


ggplot(data=final_county, aes(x=PercentMale, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Percent Male")
ggplot(data=final_county, aes(x=PercentMale, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Percent Male")

summary(lm(VOTEPERCENT~`PercentMale`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`PercentMale`, data=subset(final_county,MIDTERM==1)))
```
```{r}

ggplot(data=final_county, aes(x=PercentFemale, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Percent Female")
ggplot(data=final_county, aes(x=PercentFemale, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Percent Female")

summary(lm(VOTEPERCENT~`PercentFemale`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`PercentFemale`, data=subset(final_county,MIDTERM==1)))
```


```{r}

ggplot(data=final_county, aes(x=AGE2024_TOTAL, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Age 20 24 total")
ggplot(data=final_county, aes(x=AGE2024_TOTAL, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Age 20 24 total")

summary(lm(VOTEPERCENT~`AGE2024_TOTAL`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`AGE2024_TOTAL`, data=subset(final_county,MIDTERM==1)))
```

```{r}

ggplot(data=final_county, aes(x=AGE2024_MALE, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Age 20 24 male")
ggplot(data=final_county, aes(x=AGE2024_MALE, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Age 20 24 male")

summary(lm(VOTEPERCENT~`AGE2024_MALE`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`AGE2024_MALE`, data=subset(final_county,MIDTERM==1)))
```
```{r}

ggplot(data=final_county, aes(x=AGE2024_FEMALE, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Age 20 24 female")
ggplot(data=final_county, aes(x=AGE2024_FEMALE, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Age 20 24 female")

summary(lm(VOTEPERCENT~`AGE2024_FEMALE`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`AGE2024_FEMALE`, data=subset(final_county,MIDTERM==1)))
```

```{r}

ggplot(data=final_county, aes(x=AGE2529_TOTAL, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Age 25 29 total")
ggplot(data=final_county, aes(x=AGE2529_TOTAL, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Age 25 29 total")

summary(lm(VOTEPERCENT~`AGE2529_TOTAL`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`AGE2529_TOTAL`, data=subset(final_county,MIDTERM==1)))
```
```{r}

ggplot(data=final_county, aes(x=AGE2529_MALE, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Age 25 29 male")
ggplot(data=final_county, aes(x=AGE2529_MALE, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Age 25 29 male")

summary(lm(VOTEPERCENT~`AGE2529_MALE`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`AGE2529_MALE`, data=subset(final_county,MIDTERM==1)))
```
```{r}

ggplot(data=final_county, aes(x=AGE2529_FEMALE, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Age 25 29 female")
ggplot(data=final_county, aes(x=AGE2529_FEMALE, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Age 25 29 female")

summary(lm(VOTEPERCENT~`AGE2529_FEMALE`*MIDTERM, data = final_county))
summary(lm(VOTEPERCENT~`AGE2529_FEMALE`, data=subset(final_county,MIDTERM==1)))
```

###Interactions

```{r}
demographics <- demographics[complete.cases(demographics), ]

demographics <- demographics %>%
  mutate(highwhite = ifelse(percent_white >0.94, 1, 0), highblack = ifelse(percent_black >.007514032, 1, 0) )

```

```{r}
#final_county$INCOME.HC01_EST_VC13
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC13, y=VOTEPERCENT, color=as.factor(CTYNAME))) +
    geom_point() + facet_wrap(~YEAR) +xlab("Households; Estimate; Median income (dollars)")
ggplot(data=final_county, aes(x=INCOME.HC01_EST_VC13, y=VOTEPERCENT)) +
    geom_point() + facet_wrap(~MIDTERM) + xlab("Households; Estimate; Median income (dollars)")
#comment:positive relationship
```