---
title: "Ha_MergeData2"
author: "Thu Ha Nguyen"
date: "10/27/2018"
output: html_document
---




## Voter turnout
KEY variables: precinct name (PTCNAME), precinct code/ state precinct number (PCTCODE), county name(CTYNAME), county code (CTYCODE), voter turnout/ballot (TOTVOTERS), us congressional district (CONGDIST), year (YEAR)

#Midterm election
```{r}
#2014
turnout_pct_14.df <- read_xlsx("~/jvstats/VoterTurnout/Election_2014_Precinct.xlsx")
turnout_pct_14.df <- turnout_pct_14.df %>%
  mutate(YEAR = 2014,
         MIDTERM = 1)
colnames(turnout_pct_14.df)[c(2, 3, 5, 6,24 , 7)] <- c("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST")
turnout_pct_14.df <- type_convert(turnout_pct_14.df)
turnout_pct_14.df <- turnout_pct_14.df %>%
  mutate(PCTCODE = as.numeric(PCTCODE))

#2010
turnout_pct_10.df <- read_excel("~/jvstats/VoterTurnout/Election_2010_Precinct.xls")
turnout_pct_10.df <- turnout_pct_10.df %>%
  mutate(YEAR = 2010,
         MIDTERM = 1)
turnout_pct_10.df$County[turnout_pct_10.df$County == "SAINT LOUIS"] <- "ST. LOUIS"
colnames(turnout_pct_10.df)[c(1, 2, 10, 9, 20, 3)] <- c("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST")
turnout_pct_10.df <- turnout_pct_10.df %>%
  filter(!is.na(CTYNAME))


#2006 
#2006 doesn't have county name so we add county name in sheet 3 of the same excel file
turnout_pct_06.df <- read_excel("~/jvstats/VoterTurnout/Election_2006_Precinct.xls")
turnout_pct_06.df <- turnout_pct_06.df %>%
  filter(!is.na(County_ID), County_ID != "County_ID")
turnout_pct_06.df <- turnout_pct_06.df %>%
  mutate(CC = as.numeric(County_ID))
countyname06.df <- read_excel("~/jvstats/VoterTurnout/Election_2006_Precinct.xls", sheet = 3)
turnout_pct_06.df <- merge(turnout_pct_06.df, countyname06.df, by=c("CC"))

turnout_pct_06.df <- turnout_pct_06.df %>%
  mutate(YEAR = 2006,
         MIDTERM = 1)
colnames(turnout_pct_06.df)[c(2, 11, 64, 1, 20, 4)] <- c("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST")
turnout_pct_06.df <- turnout_pct_06.df %>%
  mutate(PCTCODE = as.numeric(PCTCODE),
         CTYCODE = as.numeric(CTYCODE),
         TOTVOTERS = as.numeric(TOTVOTERS),
         CONGDIST = as.numeric(CONGDIST))
length(unique(turnout_pct_06.df$CTYNAME))

#2002
#2002 has many variables with candidate name (?), so these columns were filtered out.
turnout_pct_02.df <- read_excel("~/jvstats/VoterTurnout/Election_2002_Precinct.xls", skip=2)
turnout_pct_02.df <- type_convert(turnout_pct_02.df)
turnout_pct_02.df <- turnout_pct_02.df %>%
  select(1:76) %>%      
  mutate(CC = as.numeric(CC))
countyname02.df <- read_excel("~/jvstats/VoterTurnout/Election_2002_Precinct.xls", sheet = 3)
turnout_pct_02.df <- merge(turnout_pct_02.df, countyname02.df, by=c("CC"))

turnout_pct_02.df <- turnout_pct_02.df %>%
  mutate(YEAR = 2002,
         MIDTERM = 1)

colnames(turnout_pct_02.df)[c(3, 17, 78, 1, 24, 5)] <- c("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST")
turnout_pct_02.df <- turnout_pct_02.df %>%
  mutate(PCTCODE = as.numeric(PCTCODE),
         CTYCODE = as.numeric(CTYCODE),
         TOTVOTERS = as.numeric(TOTVOTERS),
         CONGDIST = as.numeric(CONGDIST))
length(unique(turnout_pct_02.df$CTYNAME))

turnout_pct_14.df <- turnout_pct_14.df %>%
  select(PCTNAME, PCTCODE, CTYNAME, CTYCODE, TOTVOTERS, CONGDIST, YEAR, MIDTERM)
turnout_pct_10.df <- turnout_pct_10.df %>%
  select(PCTNAME, PCTCODE, CTYNAME, CTYCODE, TOTVOTERS, CONGDIST, YEAR, MIDTERM)
turnout_pct_06.df <- turnout_pct_06.df %>%
  select(PCTNAME, PCTCODE, CTYNAME, CTYCODE, TOTVOTERS, CONGDIST, YEAR, MIDTERM)
turnout_pct_02.df <- turnout_pct_02.df %>%
  select(PCTNAME, PCTCODE, CTYNAME, CTYCODE, TOTVOTERS, CONGDIST, YEAR, MIDTERM)


#Merge voter turnout
turnout_pct.df <- turnout_pct_14.df %>%
 rbind(turnout_pct_10.df) %>%
  rbind(turnout_pct_06.df) %>%
  rbind(turnout_pct_02.df) %>%
  select("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST", "YEAR", "MIDTERM") %>%
  mutate(PCTNAME = str_to_lower(PCTNAME),
         CTYNAME = str_to_lower(CTYNAME))
head(turnout_pct.df)
#may need to do some checking: counting the number of precinct by year/ county

turnout_cty.df <- turnout_pct.df %>%
  group_by(`CTYNAME`,`YEAR`, CTYCODE, MIDTERM) %>%
  summarise(
    TOTVOTERS = sum(TOTVOTERS))
head(turnout_cty.df)
```
#Presidential election
```{r}
#2016
turnout_pct_16.df <- read_excel("~/jvstats/PresidentialTurnout/2016-general-federal-state-results-by-precinct-official.xlsx")
turnout_pct_16.df <- turnout_pct_16.df %>%
  mutate(YEAR = 2016,
         MIDTERM = 0)
colnames(turnout_pct_16.df)[c(2, 3, 5, 6,25 , 7)] <- c("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST")
turnout_pct_16.df <- type_convert(turnout_pct_16.df)
head(turnout_pct_16.df)
turnout_pct_16.df <- turnout_pct_16.df %>%
  filter(CTYNAME != "TOTAL") %>%
  mutate(PCTCODE = as.numeric(PCTCODE))

#2012
turnout_pct_12.df <- read_excel("~/jvstats/PresidentialTurnout/2012mngeneralelectionresults_official_postrecounts.xlsx")
turnout_pct_12.df <- turnout_pct_12.df %>%
  mutate(YEAR = 2012,
         MIDTERM = 0)
colnames(turnout_pct_12.df)[c(1, 2, 4, 5,24 , 6)] <- c("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST")
turnout_pct_12.df <- type_convert(turnout_pct_12.df)
head(turnout_pct_12.df)
turnout_pct_12.df <- turnout_pct_12.df %>%
  filter(CTYNAME != "NA") %>%
  mutate(PCTCODE = as.numeric(PCTCODE))


#2008
turnout_pct_08.df <- read_excel("~/jvstats/PresidentialTurnout/2008_general_results.xls")
turnout_pct_08.df <- turnout_pct_08.df %>%
  mutate(YEAR = 2008,
         MIDTERM = 0)
ctyname08 <- read_excel("~/jvstats/PresidentialTurnout/2008_general_results.xls", sheet =3)
ctyname08 <- ctyname08 %>%
  mutate(CountyID = as.character(CountyID))
turnout_pct_08.df <- turnout_pct_08.df %>%
  left_join(ctyname08, by = c("CountyID"))

colnames(turnout_pct_08.df)[c(1, 2, 59, 14,21 , 3)] <- c("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST")
turnout_pct_08.df <- type_convert(turnout_pct_08.df)
head(turnout_pct_08.df)
turnout_pct_08.df <- turnout_pct_08.df %>%
  filter(CTYNAME != "NA", CTYNAME != "STATE ABSENTEE BALLOT") %>%
  mutate(PCTCODE = as.numeric(PCTCODE),
         CONGDIST = as.numeric(CONGDIST))

turnout_pct_16.df <- turnout_pct_16.df %>%
  select(PCTNAME, PCTCODE, CTYNAME, CTYCODE, TOTVOTERS, CONGDIST, YEAR, MIDTERM)
turnout_pct_12.df <- turnout_pct_12.df %>%
  select(PCTNAME, PCTCODE, CTYNAME, CTYCODE, TOTVOTERS, CONGDIST, YEAR, MIDTERM)
turnout_pct_08.df <- turnout_pct_08.df %>%
  select(PCTNAME, PCTCODE, CTYNAME, CTYCODE, TOTVOTERS, CONGDIST, YEAR, MIDTERM)

#Merge presidential election
turnout_pct_pre.df <- turnout_pct_16.df %>%
  rbind(turnout_pct_12.df) %>%
  rbind(turnout_pct_08.df) %>%
  select("PCTNAME", "PCTCODE", "CTYNAME", "CTYCODE", "TOTVOTERS", "CONGDIST", "YEAR", "MIDTERM") %>%
  mutate(PCTNAME = str_to_lower(PCTNAME),
         CTYNAME = str_to_lower(CTYNAME))
head(turnout_pct_pre.df)

turnout_cty_pre.df <- turnout_pct_pre.df %>%
  group_by(`CTYNAME`,`YEAR`, CTYCODE, MIDTERM) %>%
  summarise(
    TOTVOTERS = sum(TOTVOTERS))
head(turnout_cty_pre.df)
head(turnout_cty.df)
```
#Merge midterm and presidential election
```{r}
# !diagnostics off
sum(is.na(turnout_pct_all.df$TOTVOTERS))

turnout_cty_all.df <- rbind(turnout_cty.df, turnout_cty_pre.df)
turnout_pct_all.df <- rbind(turnout_pct.df, turnout_pct_pre.df)
#write_excel_csv(turnout_cty_all.df, "~/jvstats/NEW_DATA/EXPORTED_DATA/turnout_cty.csv")
#write_excel_csv(turnout_pct_all.df, "~/jvstats/NEW_DATA/EXPORTED_DATA/turnout_pct.csv")
```
#Political variables

## Cook's political report

```{r}
Cook.df <- read_csv("~/jvstats/CookCompetitiveness.csv")
head(Cook.df)
Cook1.df <- Cook.df %>%
  select(-ReportDate) %>%
  filter(OfficeName == "House Representative") %>%
  mutate(HouseRep = Competitiveness) %>%
  select(- OfficeName, - Competitiveness)

Cook2.df <- Cook.df %>%
  select(-ReportDate, - District) %>%
  filter(OfficeName != "House Representative") %>%
  spread(key = OfficeName, value = Competitiveness)

Cook1.df <- rbind(Cook1.df, c(2014, 2, 0), c(2014, 3, 0), c(2014, 4, 0), c(2014, 5, 0), c(2014, 6, 0))
Cook1.df <- rbind(Cook1.df, c(2018, 4, 0), c(2018, 5, 0), c(2018, 6, 0))

Cook.df <- Cook1.df %>%
  left_join(Cook2.df)
head(Cook1.df)
head(Cook2.df)
colnames(Cook.df) <- c("YEAR", "CONGDIST", "COOK.HouseRep", "COOK.Governer", "COOK.Senate1", "COOK.Senate2")
Cook.df <- Cook.df %>%
  mutate(COOK.HouseRep = as.numeric(COOK.HouseRep),
         COOK.Governer = as.numeric(COOK.Governer),
         COOK.Senate1 = as.numeric(COOK.Senate1),
         COOK.Senate2 = as.numeric(COOK.Senate2))
```

#Social capital
```{r}
# !diagnostics off

socialcap_14.df <- read_excel("~/jvstats/NEW_DATA/POLITICAL_VARIABLES/2014_SocialCapital.xlsx")
colnames(socialcap_14.df) <- str_c("SOCIALCAP.", colnames(socialcap_14.df)) 
socialcap_14.df <- socialcap_14.df %>%
  filter(str_sub(SOCIALCAP.County_Name, -2, -1) == "MN") %>%
  mutate(CTYNAME = str_to_lower(str_sub(SOCIALCAP.County_Name, 1 , str_length(SOCIALCAP.County_Name) - 11))) %>%
  select(-SOCIALCAP.County_Name, -SOCIALCAP.FIPS) %>%
    mutate(CTYNAME = as.character(CTYNAME))
head(socialcap_14.df)


socialcap_09.df <- read_excel("~/jvstats/NEW_DATA/POLITICAL_VARIABLES/1997_2005_2009_SocialCapital.xlsx", sheet = 3)
colnames(socialcap_09.df) <- str_c("SOCIALCAP.", colnames(socialcap_09.df)) 
socialcap_09.df <- socialcap_09.df %>%
  filter(str_sub(SOCIALCAP.areaname, -2, -1) == "MN") %>%
  mutate(CTYNAME = str_to_lower(str_sub(SOCIALCAP.areaname, 1 , str_length(SOCIALCAP.areaname) - 4))) %>%
  select(-SOCIALCAP.areaname, -SOCIALCAP.fips) %>%
  mutate(CTYNAME = as.character(CTYNAME))
head(socialcap_09.df)

#merge social capital variable
socialcap.df <- cbind(socialcap_14.df, socialcap_09.df)

#write_excel_csv(socialcap.df, "~/jvstats/NEW_DATA/EXPORTED_DATA/socialcap.csv")

```

#House control
```{r}
housecontrol.df <- read_csv("~/jvstats/NEW_DATA/PARTISAN_CONTROL/CONTROL_HOUSE.csv")
head(housecontrol.df)
colnames(housecontrol.df) <- str_c("HOUSE.", colnames(housecontrol.df))
housecontrol.df <- housecontrol.df %>%
  mutate(YEAR = (as.numeric(str_sub(`HOUSE.Congress (Years)`, -5,-2)) + as.numeric(str_sub(`HOUSE.Congress (Years)`, -10,-7)))/2)
head(housecontrol.df)
```

#Senate control
```{r}
senatecontrol.df <- read_csv("~/jvstats/NEW_DATA/PARTISAN_CONTROL/CONTROL_SENATE.csv")
head(senatecontrol.df)
colnames(senatecontrol.df) <- str_c("SENATE.", colnames(senatecontrol.df))
head(senatecontrol.df)
senatecontrol.df <- senatecontrol.df %>%
  mutate(YEAR = (as.numeric(str_sub(`SENATE.Session`, -5,-2)) + as.numeric(str_sub(`SENATE.Session`, -10,-7)))/2)
head(senatecontrol.df)
```

#State legislative
```{r}
statelegis.df <- read_csv("~/jvstats/NEW_DATA/PARTISAN_CONTROL/CONTROL_STATE_LEGISLATURE.csv")
statelegis.df <- statelegis.df[c(1:9),]
colnames(statelegis.df) <- str_c("STATELEGIS.", colnames(statelegis.df))

statelegis.df <- statelegis.df %>%
  mutate(YEAR = as.numeric(str_sub(`STATELEGIS.Session`, -4,-1)))
head(statelegis.df)
```


#Merge everything
```{r}
#main.pct <- turnout_pct_all.df %>%
  #left_join(income_cty.df, by = c("YEAR", "CTYNAME")) %>%  #use code book of 2014
#  left_join(socialcap.df, by = c("CTYNAME")) %>%
#  left_join(housecontrol.df, by = c("YEAR")) %>%
#  left_join(senatecontrol.df, by = c("YEAR")) %>%
#  left_join(statelegis.df, by = c("YEAR"))
#main.pct <- main.pct %>%
#  left_join(Cook.df, by = c("YEAR", "CONGDIST"))
#head(main.pct)
#write_csv(main.df, "sampledata.csv")


socialcap <- read_csv("~/jvstats/NEW_DATA/EXPORTED_DATA/socialcap.csv",
                      col_types = cols(
                              CTYNAME_1 = col_character()
                            ))
socialcap <- socialcap %>%
  mutate(CTYNAME = as.character(CTYNAME_1))
main.cty1 <- turnout_cty_all.df %>%
  left_join(socialcap, by = c("CTYNAME"))

main.cty <- main.cty1 %>%
  left_join(housecontrol.df, by=c("YEAR")) %>%
  left_join(senatecontrol.df, by=c("YEAR")) %>%
  left_join(statelegis.df, by=c("YEAR")) %>%
  select(-CTYNAME_1)
head(main.cty)
#write_excel_csv(main.cty, "~/jvstats/NEW_DATA/EXPORTED_DATA/turnout_political_cty.csv")
```
