---
title: "MYKHE_Data_2018"
author: "My Khe"
date: "10/31/2018"
output: html_document
---

---
title: "Ha_DemographicTrend"
author: "Thu Ha Nguyen"
date: "10/30/2018"
output: html_document
---

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
```


```{r}
final_county <- read_csv("~/jvstats/final_county.csv")
#summary(final_county)
```
#voting age eligible population

```{r}
ggplot(data=final_county, aes(x=YEAR, y=VOTE_ELIGIBLE, color=as.factor(CTYNAME))) +
    geom_line()

#https://metrocouncil.org/News-Events/Communities/News-Articles/Population-growth-in-the-7-county-metro-remains-st.aspx

final_county[which.max(final_county$VOTE_ELIGIBLE),]
CTYNAME.vector <- as.data.frame(final_county[final_county$YEAR==2016, "CTYNAME"])
VOTE_ELIGIBLE.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(VOTE_ELIGIBLE ~ YEAR, data=subdata))
  VOTE_ELIGIBLE.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

VOTE_ELIGIBLE.vector

data_2018 <- cbind(CTYNAME.vector, VOTE_ELIGIBLE.vector)
data_2018
```

```{r}
ggplot(data=final_county, aes(x=YEAR, y=`Citizen: US Born`, color=as.factor(CTYNAME))) +
    geom_line() 
CITIZEN.USBORN.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`Citizen: US Born` ~ YEAR, data=subdata))
  CITIZEN.USBORN.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

CITIZEN.USBORN.vector

data_2018 <- cbind(data_2018, CITIZEN.USBORN.vector)

ggplot(data=final_county, aes(x=YEAR, y=`Citizen: Born Abroad`, color=as.factor(CTYNAME))) +
    geom_line() 
CITIZEN.BORNABROAD.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`Citizen: Born Abroad` ~ YEAR, data=subdata))
  CITIZEN.BORNABROAD.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

CITIZEN.BORNABROAD.vector

data_2018 <- cbind(data_2018, CITIZEN.BORNABROAD.vector)
data_2018

```

```{r}
# Agriculture
final_county$OCC.HC01_EST_VC02 <- as.numeric(final_county$OCC.HC01_EST_VC02)
ggplot(data=final_county, aes(x=YEAR, y=OCC.HC01_EST_VC02, color=as.factor(CTYNAME))) +
    geom_line() 
ggplot(data=final_county[final_county$CTYNAME == "ramsey",], aes(x=YEAR, y=OCC.HC01_EST_VC02, color=as.factor(CTYNAME))) +
    geom_line() 

ggplot(data=final_county[final_county$CTYNAME == "redwood",], aes(x=YEAR, y=OCC.HC01_EST_VC02, color=as.factor(CTYNAME))) +
    geom_line() 

ggplot(data=final_county[final_county$CTYNAME == "red lake",], aes(x=YEAR, y=OCC.HC01_EST_VC02, color=as.factor(CTYNAME))) +
    geom_line() 

CITIZEN.USBORN.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`Citizen: US Born` ~ YEAR, data=subdata))
  CITIZEN.USBORN.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

CITIZEN.USBORN.vector

data_2018 <- cbind(data_2018, CITIZEN.USBORN.vector)

ggplot(data=final_county, aes(x=YEAR, y=`Citizen: Born Abroad`, color=as.factor(CTYNAME))) +
    geom_line() 
CITIZEN.BORNABROAD.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`Citizen: Born Abroad` ~ YEAR, data=subdata))
  CITIZEN.BORNABROAD.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

CITIZEN.BORNABROAD.vector

data_2018 <- cbind(data_2018, CITIZEN.BORNABROAD.vector)
data_2018
```
#employment and income
```{r}
ggplot(data=final_county, aes(x=YEAR, y=EMPLOY.HC04_EST_VC01, color=as.factor(CTYNAME))) +
    geom_line() + ylab("Unemployed; Estimate; Population 16 years and over")

ggplot(data=final_county, aes(x=YEAR, y=INCOME.HC01_EST_VC13, color=as.factor(CTYNAME))) +
    geom_line() + ylab("Households; Estimate; Median income (dollars)")

ggplot(data=final_county, aes(x=YEAR, y=INCOME.HC01_EST_VC02, color=as.factor(CTYNAME))) +
    geom_line() + ylab("Households; Estimate; Less than $10,000")
summary(lm(INCOME.HC01_EST_VC13~YEAR, data=final_county))
```

#education
```{r}
#%HS Degree
Education_hs.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`HIGHSCHOOL DEGREE` ~ YEAR, data=subdata))
  Education_hs.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

Education_hs.vector

data_2018 <- cbind(data_2018, Education_hs.vector)

ggplot(data=final_county, aes(x=YEAR, y=`HIGHSCHOOL DEGREE`, color=as.factor(CTYNAME))) +
    geom_line() 

```

```{r}
#%bachelor's
Education_ba.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`BACHELOR'S DEGREE` ~ YEAR, data=subdata))
  Education_ba.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

Education_ba.vector

data_2018 <- cbind(data_2018, Education_ba.vector)

ggplot(data=final_county, aes(x=YEAR, y=`BACHELOR'S DEGREE`, color=as.factor(CTYNAME))) +
    geom_line() 

#note cook county is an outlier
ba_outlier<-education.df%>%
filter(CTYNAME=="Cook County, Minnesota")

  ggplot(data=ba_outlier, aes(x=YEAR, y=`BACHELOR'S DEGREE`)) +
    geom_line()+
    geom_smooth(method=lm)
```


# Income (Median Household Income, Percent of Households that have annual income less than $10,000), 
```{r}
#Median Income
#INCOME.HC01_EST_VC13
Median_income.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`INCOME.HC01_EST_VC13` ~ YEAR, data=subdata))
  Median_income.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

Median_income.vector

data_2018 <- cbind(data_2018, Median_income.vector)

ggplot(data=final_county, aes(x=YEAR, y=`INCOME.HC01_EST_VC13`, color=as.factor(CTYNAME))) +
    geom_line()+
  labs(title="Regression of Median Income by year",y="median income", colour = "County")
```

```{r}
#Househols<10,000
#INCOME.HC01_EST_VC02
low_income.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`INCOME.HC01_EST_VC02` ~ YEAR, data=subdata))
  low_income.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

low_income.vector

data_2018 <- cbind(data_2018, low_income.vector)

ggplot(data=final_county, aes(x=YEAR, y=`INCOME.HC01_EST_VC02`, color=as.factor(CTYNAME))) +
    geom_line()+
  labs(title="Regression of Low Income Households by year",y="%Low Income", colour = "County")

```

#Age(Percent of the population 20-24 years old, percent of the population 25-29 years old)
```{r}
#20-24
#AGE2024
age20_24.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`AGE2024` ~ YEAR, data=subdata))
  age20_24.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

age20_24.vector

data_2018 <- cbind(data_2018, age20_24.vector)

ggplot(data=final_county, aes(x=YEAR, y=`AGE2024`, color=as.factor(CTYNAME))) +
    geom_line()+
  labs(title="Regression Age 20-24",y="%20-24", colour = "County")

#note: residuals may be producing problematic data

```

```{r}
#20-24
#AGE2529
age25_29.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`AGE2529` ~ YEAR, data=subdata))
  age25_29.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

age25_29.vector

data_2018 <- cbind(data_2018, age20_24.vector)

ggplot(data=final_county, aes(x=YEAR, y=`AGE2529`, color=as.factor(CTYNAME))) +
    geom_line()+
  labs(title="Regression Age 25-29",y="%25-29", colour = "County")

```

#Race(Percent of the population of white, Percent of the population of black)
```{r}
#percent white
#define variable for final data
final_county<- final_county%>%
  group_by(CTYNAME)%>%
  mutate(percent_white=WHITE/POPULATION)


percentwhite.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(percent_white ~ YEAR, data=subdata))
  percentwhite.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

percentwhite.vector

data_2018 <- cbind(data_2018, age20_24.vector)

ggplot(data=final_county, aes(x=YEAR, y=`percent_white`, color=as.factor(CTYNAME))) +
    geom_line()+
  labs(title="Regression of white population",y="% white", colour = "County")

```
```{r}
#percent black
#define variable for final data
final_county<- final_county%>%
  group_by(CTYNAME)%>%
  mutate(percent_black=BLACK/POPULATION)


percentblack.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(percent_black ~ YEAR, data=subdata))
  percentblack.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

percentblack.vector

data_2018 <- cbind(data_2018, percentblack.vector)

ggplot(data=final_county, aes(x=YEAR, y=`percent_black`, color=as.factor(CTYNAME))) +
    geom_line()+
  labs(title="Regression of black population",y="% black", colour = "County")

```

```{r}
#percent unemployed

unemployment_rate.vector <- NULL
for (i in 1:nrow(CTYNAME.vector)) {
  subdata <- final_county[final_county$CTYNAME == CTYNAME.vector[[i,1]], ]
  model <- summary(lm(`EMPLOY.HC04_EST_VC01` ~ YEAR, data=subdata))
  unemployment_rate.vector[i] <- model$coefficients[1] + model$coefficients[2] * 2018
}

unemployment_rate.vector

data_2018 <- cbind(data_2018, unemployment_rate.vector)

ggplot(data=final_county, aes(x=YEAR, y=`EMPLOY.HC04_EST_VC01`, color=as.factor(CTYNAME))) +
    geom_line()+
  labs(title="Regression of unemplyment rate",y="%unemployed", colour = "County")


```

